### 2025-02-10 Manuscript.qmd
Update method (before data analysis)

---
title: "Exploring the Impact of Queer Media Characters on LGBTQ+ Psychological Well-Being: The Role of Queer Media Consumption, Perceived Authenticity, and Intersectionality"
shorttitle: "Queer Media Characters and LGBTQ+ Well-Being"
author:
  - name: Dingning Yang
    corresponding: true
    orcid: 0009-0008-5546-962X
    email: dingningy@uchicago.edu
    affiliations:
      - name: University of Chicago
        department: Department of Social Sciences
        address: 5105 South Harper Ave.
        city: Chicago
        region: IL
        country: USA
        postal-code: 60615-4156
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: |

  | The growing visibility of the LGBTQ+ community is accompanied by the flourishing of queer characters in media in the information era. While previous studies have explored the double-edged impact of queer media representation on LGBTQ+ individuals’ development, few have compared these effects across different LGBTQ+ subgroups. This study investigates queer media characters’ psychological impacts on LGBTQ+ individuals with different identities, examining the relationship between perceived character authenticity, intersectionality, and characters’ impacts on LGBTQ+ individuals’ psychological well-being using a reliable survey instrument in a worldwide sample.

  | Findings revealed that marginal groups within the LGBTQ+ community, including those from developing countries, gender minorities, and racial minorities, reported higher scores about queer media characters’ positive impacts on psychological well-being compared to their mainstream counterparts. However, in terms of sexual orientation, mainstream groups (i.e., gay and lesbian respondents) showed higher scores. Importantly, the study found strong positive correlations between authenticity ratings, intersectionality ratings, and scores of queer characters’ positive impacts on psychological well-being.

  | This study contributes to a more nuanced understanding of LGBTQ+ individuals' diverse needs and experiences with queer media representation, which has significant implications for media producers, mental health professionals, and LGBTQ+ advocacy groups in fostering more positive and inclusive media representations for the LGBTQ+ community.
keywords: [LGBTQ+ well-being, queer media representations, intersectionality]
format:
  apaquarto-pdf:
    csl: apa.csl
    documentmode: man
    indent: true
    include-in-header: 
      text: |
        \usepackage{setspace}
        \linespread{2}
        \renewcommand{\footnotesize}{\fontsize{8}{10}\selectfont}
bibliography: References.bib
---
Since the Stonewall riots, queer characters have flourished in the mass media, reaching their peak in the information era. As the primary audience for queer media characters, the LGBTQ+ community is claimed to be both positively and negatively impacted by these characters [e.g., @craig_media:_2015;@nolke_making_2018]. While queer characters can favorably impact LGBTQ+ individuals’ identity development and self-esteem, stereotypical representations that lack intersectionality may result in stigmatization and negative mental outcomes. Nevertheless, few studies have compared queer characters’ impacts on different LGBTQ+ subgroups, probably because they ignored the emergence of more marginal groups (e.g., asexual). Additionally, no established instruments have explored how perceived character authenticity or intersectionality may contribute to different psychological well-being outcomes. Knowing LGBTQ+ individuals' diverse needs and experiences with queer media representation is important as it can call for more diverse and inclusive queer media representations. Through developing a reliable instrument on queer media characters’ psychological impacts and conducting worldwide data collection across different LGBTQ+ generations, this study aimed to contribute to a deeper understanding of queer characters’ psychological impacts on different LGBTQ+ groups.

# Literature Review

## LGBTQ+ Community and Queer Media Characters

@IpsosLimited2024 has found that 17% of Generation Zers^[Generation Z (born between 1996-2012), Millennials (born between 1980-1995), Generation X (born between 1966-1979) and Baby Boomers (born between 1945-1965).] identified as part of the LGBTQ+ community, including identities such as lesbian, gay, bisexual, transgender, queer, asexual, and non-binary. This indicates a significant increase compared with prior generations, with only 11% for Millennials, 6% for Generation Xers, and 5% for Baby Boomers. The evolution within the LGBTQ+ community is accompanied by the flourishing of queer characters in TV shows and films that cater to their needs for identity affirmation and psychological belonging. For instance, the past ten years have witnessed a more than 300% increase in the number of LGBTQ+ characters on broadcast [@GLAAD2024]. Compared with earlier queer characters that often feature heavy stereotypes (e.g., Frank-N-Furter in *Rocky Horror*), queer characters nowadays feature more multidimensional personalities and diversity, indicating progress in the representation of queer lives in media.

## Queer Media Characters’ Positive Impacts on the LGBTQ+ Community

@toomey_coping_2018 indicated that LGBTQ+ individuals may rely on social and cultural resources to enhance identity affirmation and cope with stress. Prior studies found that queer media characters can positively affect LGBTQ+ individuals as an effective coping strategy. @gomillion_influence_2011 conducted surveys and interviews on lesbian, gay, and bisexual (LGB) youth from Texas, suggesting that queer media characters, as role models, could positively affect LGB youth’s self-identity by enhancing their self-esteem and courage. While @gomillion_influence_2011 ignored transgenders and genderqueers, @craig_media:_2015 showed that queer media characters could help LGBTQ youth improve their self-assurance and build relationships with people having shared identities. Nevertheless, prior studies convergently focused on queer characters’ impacts on LGBTQ+ youth in the United States, ignoring cross-generation and cross-country comparisons, which may result in low external validity of the findings. Additionally, these studies barely compared different sexual orientations and gender identities in the LGBTQ+ community, which may neglect more marginalized groups’ perceptions and experiences (e.g., asexual and pansexual).

## Queer Media Stereotypes’ Negative Impacts on the LGBTQ+ Community

Some studies have also criticized that queer media characters presenting stereotypes could be harmful to the LGBTQ+ community. @yaksich_consuming_2005 and @guidotto_cashing_2006 indicated that many queer media representations are the commodification of the LGBTQ+ community. The queer media monopoly of ‘perfect homo’ (i.e., flamboyant Caucasian and middle-classed gay men) neglects intersectionality, reinforcing the LGBTQ+ stereotypes in the audience’s perceptions. @nolke_making_2018 identified the intersectional invisibility of queer media representations by applying intersectional analysis of a series of advertisements from 2009 to 2015. The discussion from this study also indicated that the hegemony of queer stereotypes and the lack of intersectionality could be harmful to the LGBTQ+ community by bringing them lower confidence and mental illnesses (e.g., depressive disorders).
Nevertheless, no established instrument has systematically compared how LGBTQ+ perceived queer character authenticity (i.e., free of stereotypes) or perceived queer character intersectionality could impact their psychological well-being, calling for more scholarly intention to relevant measurement.

# Current Study and Hypotheses

The current study examined queer media characters’ impacts on LGBTQ+ individuals’ psychological well-being through two overarching research questions. The first research question investigated whether there were differences in queer characters’ psychological impacts among LGBTQ+ individuals with different identities. Specifically, this study has compared LGBTQ+ individuals with different sexual orientations, gender identities, races, ages, and nationalities on queer characters’ psychological impacts. Given the fact that there were still queer stereotypes and fewer intersectional representations in mass media, the current study stated that queer media characters’ positive impacts on psychological well-being would be weaker for LGBTQ+ individuals with more invisible and marginalized identities (e.g., asexual, transgender, black, elder generations, developing countries’ residents) while being stronger for those with more visible identities (e.g., gay, cisgender, white, younger generations, developed countries’ residents). The second research question asked whether queer media consumption patterns, perceived character authenticity, and perceived character intersectionality can predict queer media characters’ psychological impacts. Specifically, this study hypothesized that lower ratings for perceived character authenticity and intersectionality would result in lower ratings for queer characters’ positive psychological impacts. Conversely, higher ratings for perceived character authenticity and intersectionality would result in higher ratings for queer characters’ positive psychological impacts.

# Method

## Participants
This study recruited 180 LGBTQ+ participants worldwide through the data collection platform Prolific. Specifically, the respondents were aged from 18 to 72 in 17 countries (including countries in North America, Europe, Asia, and Africa) with different sexual orientations, gender identities, and racial identities, satisfying the research needs for comparisons among LGBTQ+ individuals with different identities on queer media characters’ psychological impacts. Each participant was paid $2 to participate in a survey with a seven-minute average finish time.

## Measurements

This study developed a survey instrument on Qualtrics to explore queer media characters’ psychological impacts among different LGBTQ+ individuals with five sections:

* ***Demographic form***: The demographic form includes nine questions that ask participants’ age, residential country (and state if living in the US), gender identity, sexual orientation, racial identity, and education levels.

* ***Queer media consumption***: The queer media consumption scale asks five questions regarding participants’ exposure and engagement with queer media characters. Specifically, the scale measures the frequency of watching queer TV shows or movies, the discussion engagement of queer representations on social media, and the specific queer media genre and names as a mix of close-ended and open-ended questions.

* ***Queer character authenticity***: The queer character authenticity scale consists of seven questions, with four of them asking the extent to which queer characters can reflect individual experiences in daily life. The other three questions measure the extent to which queer characters displaying stereotypes, which are reverse-coded. The sum of seven questions consists of the authenticity rating of the whole scale.

* ***Queer character intersectionality***: The queer character intersectionality scale consists of four questions that ask the extent to which queer characters can reflect diverse experiences of multiple factors such as racial and socioeconomic backgrounds.

* ***Queer character’s psychological impacts***: This scale measures queer characters’ impacts on participants’ psychological well-being, which is constituted of three sub-scales: 
  + Self-development (seven questions)
  + Community belonging (six questions)
  + Emotional well-being (eight questions)

# Results

# Discussion

## Limitations and Future Directions

## Conclusion

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

# Appendix

# Title for Appendix