### 2025-02-11 Manuscript.qmd
Update results of hypo 2

---
title: "Exploring the Impact of Queer Media Characters on LGBTQ+ Psychological Well-Being: The Role of Queer Media Consumption, Perceived Authenticity, and Intersectionality"
shorttitle: "Queer Media Characters and LGBTQ+ Well-Being"
author:
  - name: Dingning Yang
    corresponding: true
    orcid: 0009-0008-5546-962X
    email: dingningy@uchicago.edu
    affiliations:
      - name: University of Chicago
        department: Department of Social Sciences
        address: 5105 South Harper Ave.
        city: Chicago
        region: IL
        country: USA
        postal-code: 60615-4156
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: |

  | The growing visibility of the LGBTQ+ community is accompanied by the flourishing of queer characters in media in the information era. While previous studies have explored the double-edged impact of queer media representation on LGBTQ+ individuals’ development, few have compared these effects across different LGBTQ+ subgroups. This study investigates queer media characters’ psychological impacts on LGBTQ+ individuals with different identities, examining the relationship between perceived character authenticity, intersectionality, and characters’ impacts on LGBTQ+ individuals’ psychological well-being using a reliable survey instrument in a worldwide sample.

  | Findings revealed that marginal groups within the LGBTQ+ community, including those from developing countries, gender minorities, and racial minorities, reported higher scores about queer media characters’ positive impacts on psychological well-being compared to their mainstream counterparts. However, in terms of sexual orientation, mainstream groups (i.e., gay and lesbian respondents) showed higher scores. Importantly, the study found strong positive correlations between authenticity ratings, intersectionality ratings, and scores of queer characters’ positive impacts on psychological well-being.

  | This study contributes to a more nuanced understanding of LGBTQ+ individuals' diverse needs and experiences with queer media representation, which has significant implications for media producers, mental health professionals, and LGBTQ+ advocacy groups in fostering more positive and inclusive media representations for the LGBTQ+ community.
keywords: [LGBTQ+ well-being, queer media representations, intersectionality]
floatsintext: true
format:
  apaquarto-pdf:
    csl: apa.csl
    documentmode: man
    indent: true
    include-in-header: 
      text: |
        \usepackage{setspace}
        \linespread{2}
        \renewcommand{\footnotesize}{\fontsize{8}{10}\selectfont}
bibliography: References.bib
---

```{r}
#| label: setup
#| include: false
# Install required packages
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(psych)){install.packages("psych")}
if(!require(knitr)){install.packages("knitr")}
if(!require(kableExtra)){install.packages("kableExtra")}

# List of required packages
required_packages <- c ("tidyverse", "ggplot2", "psych","knitr", "kableExtra")

# Load required packages
lapply(required_packages, library, character.only = TRUE)
set.seed(123)

```

Since the Stonewall riots, queer characters have flourished in the mass media, reaching their peak in the information era. As the primary audience for queer media characters, the LGBTQ+ community is claimed to be both positively and negatively impacted by these characters [e.g., @craig_media:_2015;@nolke_making_2018]. While queer characters can favorably impact LGBTQ+ individuals’ identity development and self-esteem, stereotypical representations that lack intersectionality may result in stigmatization and negative mental outcomes. Nevertheless, few studies have compared queer characters’ impacts on different LGBTQ+ subgroups, probably because they ignored the emergence of more marginal groups (e.g., asexual). Additionally, no established instruments have explored how perceived character authenticity or intersectionality may contribute to different psychological well-being outcomes. Knowing LGBTQ+ individuals' diverse needs and experiences with queer media representation is important as it can call for more diverse and inclusive queer media representations. Through developing a reliable instrument on queer media characters’ psychological impacts and conducting worldwide data collection across different LGBTQ+ generations, this study aimed to contribute to a deeper understanding of queer characters’ psychological impacts on different LGBTQ+ groups.

# Literature Review

## LGBTQ+ Community and Queer Media Characters

@IpsosLimited2024 has found that 17% of Generation Zers^[Generation Z (born between 1996-2012), Millennials (born between 1980-1995), Generation X (born between 1966-1979) and Baby Boomers (born between 1945-1965).] identified as part of the LGBTQ+ community, including identities such as lesbian, gay, bisexual, transgender, queer, asexual, and non-binary. This indicates a significant increase compared with prior generations, with only 11% for Millennials, 6% for Generation Xers, and 5% for Baby Boomers. The evolution within the LGBTQ+ community is accompanied by the flourishing of queer characters in TV shows and films that cater to their needs for identity affirmation and psychological belonging. For instance, the past ten years have witnessed a more than 300% increase in the number of LGBTQ+ characters on broadcast [@GLAAD2024]. Compared with earlier queer characters that often feature heavy stereotypes (e.g., Frank-N-Furter in *Rocky Horror*), queer characters nowadays feature more multidimensional personalities and diversity, indicating progress in the representation of queer lives in media.

## Queer Media Characters’ Positive Impacts on the LGBTQ+ Community

@toomey_coping_2018 indicated that LGBTQ+ individuals may rely on social and cultural resources to enhance identity affirmation and cope with stress. Prior studies found that queer media characters can positively affect LGBTQ+ individuals as an effective coping strategy. @gomillion_influence_2011 conducted surveys and interviews on lesbian, gay, and bisexual (LGB) youth from Texas, suggesting that queer media characters, as role models, could positively affect LGB youth’s self-identity by enhancing their self-esteem and courage. While @gomillion_influence_2011 ignored transgenders and genderqueers, @craig_media:_2015 showed that queer media characters could help LGBTQ youth improve their self-assurance and build relationships with people having shared identities. Nevertheless, prior studies convergently focused on queer characters’ impacts on LGBTQ+ youth in the United States, ignoring cross-generation and cross-country comparisons, which may result in low external validity of the findings. Additionally, these studies barely compared different sexual orientations and gender identities in the LGBTQ+ community, which may neglect more marginalized groups’ perceptions and experiences (e.g., asexual and pansexual).

## Queer Media Stereotypes’ Negative Impacts on the LGBTQ+ Community

Some studies have also criticized that queer media characters presenting stereotypes could be harmful to the LGBTQ+ community. @yaksich_consuming_2005 and @guidotto_cashing_2006 indicated that many queer media representations are the commodification of the LGBTQ+ community. The queer media monopoly of ‘perfect homo’ (i.e., flamboyant Caucasian and middle-classed gay men) neglects intersectionality, reinforcing the LGBTQ+ stereotypes in the audience’s perceptions. @nolke_making_2018 identified the intersectional invisibility of queer media representations by applying intersectional analysis of a series of advertisements from 2009 to 2015. The discussion from this study also indicated that the hegemony of queer stereotypes and the lack of intersectionality could be harmful to the LGBTQ+ community by bringing them lower confidence and mental illnesses (e.g., depressive disorders).
Nevertheless, no established instrument has systematically compared how LGBTQ+ perceived queer character authenticity (i.e., free of stereotypes) or perceived queer character intersectionality could impact their psychological well-being, calling for more scholarly intention to relevant measurement.

# Current Study and Hypotheses

The current study examined queer media characters’ impacts on LGBTQ+ individuals’ psychological well-being through two overarching research questions. The first research question investigated whether there were differences in queer characters’ psychological impacts among LGBTQ+ individuals with different identities. Specifically, this study has compared LGBTQ+ individuals with different sexual orientations, gender identities, races, ages, and residential countries on queer characters’ psychological impacts. Given the fact that there were still queer stereotypes and fewer intersectional representations in mass media, the current study stated that queer media characters’ positive impacts on psychological well-being would be weaker for LGBTQ+ individuals with more invisible and marginalized identities (e.g., asexual, transgender, black, elder generations, developing countries’ residents) while being stronger for those with more visible identities (e.g., gay, cisgender, white, younger generations, developed countries’ residents). The second research question asked whether queer media consumption patterns, perceived character authenticity, and perceived character intersectionality can predict queer media characters’ psychological impacts. Specifically, this study hypothesized that lower ratings for perceived character authenticity and intersectionality would result in lower ratings for queer characters’ positive psychological impacts. Conversely, higher ratings for perceived character authenticity and intersectionality would result in higher ratings for queer characters’ positive psychological impacts.

# Method

## Participants
This study recruited 180 LGBTQ+ participants worldwide through the data collection platform Prolific. Specifically, the respondents were aged from 18 to 72 in 17 countries (including countries in North America, Europe, Asia, and Africa) with different sexual orientations, gender identities, and racial identities, satisfying the research needs for comparisons among LGBTQ+ individuals with different identities on queer media characters’ psychological impacts. Each participant was paid $2 to participate in a survey with a seven-minute average finish time.

## Measurements

This study developed a survey instrument on Qualtrics to explore queer media characters’ psychological impacts among different LGBTQ+ individuals with five sections:

* ***Demographic form***: The demographic form includes nine questions that ask participants’ age, residential country (and state if living in the US), gender identity, sexual orientation, racial identity, and education levels.

* ***Queer media consumption***: The queer media consumption scale asks five questions regarding participants’ exposure and engagement with queer media characters. Specifically, the scale measures the frequency of watching queer TV shows or movies, the discussion engagement of queer representations on social media, and the specific queer media genre and names as a mix of close-ended and open-ended questions.

* ***Queer character authenticity***: The queer character authenticity scale consists of seven questions, with four of them asking the extent to which queer characters can reflect individual experiences in daily life. The other three questions measure the extent to which queer characters displaying stereotypes, which are reverse-coded. The sum of seven questions consists of the authenticity rating of the whole scale.

* ***Queer character intersectionality***: The queer character intersectionality scale consists of four questions that ask the extent to which queer characters can reflect diverse experiences of multiple factors such as racial and socioeconomic backgrounds.

* ***Queer character’s psychological impacts***: This scale measures queer characters’ impacts on participants’ psychological well-being, which is constituted of three sub-scales: 
  + Self-development (seven questions)
  + Community belonging (six questions)
  + Emotional well-being (eight questions)

## Data Analysis

```{r}
#| label: read in cleaned and transformed datasets
#| include: false
# Read in the cleaned and transformed dataset
queer.cleaned <- read_csv("data/queer.cleaned.csv")
queer.hypo <- read_csv("data/queer.hypo.csv")
media.long <- read_csv("data/media.long.csv")

```

Data was preprocessed through RStudio and transformed into a tidy format for analysis.

```{r}
#| label: Source data analysis script
#| include: false
source("Source/Data Analysis.R")

```

### Reliability Tests
Reliability analyses were conducted to assess the internal consistency of the measurement scales used in this study (see Table 1). Cronbach's alpha coefficients were calculated for each scale. Specifically, the authenticity rating measurement demonstrated satisfactory reliability (α = 0.76). The intersectionality rating measurement showed excellent reliability (α = 0.91). Both the media individual development scale and the media community belonging scale exhibited excellent reliability (α = 0.96 for both). The emotional well-being measurement also demonstrated excellent reliability (α = 0.96). Finally, the overall psychological well-being measurement showed the highest reliability with an excellent Cronbach's alpha of 0.98. These results indicate that all scales used in this study have good to excellent internal consistency, suggesting they are reliable measures of their respective constructs.

```{r}
#| label: Generate reliability test tables
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Reliability Analysis Results"

reliability_kable <- data.frame(
  Scale = c("Authenticity rating", 
            "Intersectionality rating", 
            "Queer media individual development impact", 
            "Queer media community belonging impact", 
            "Queer media Emotional well-being impact", 
            "Queer media overall psychological well-being impact"),
  Items = c("7", "4", "7", "6", "8", "21"),
  Alpha = c(
    psych::alpha(queer.construct[,20:26], check.keys=TRUE)$total$raw_alpha,
    psych::alpha(queer.construct[,27:30])$total$raw_alpha,
    psych::alpha(queer.construct[,31:37])$total$raw_alpha,
    psych::alpha(queer.construct[,38:43])$total$raw_alpha,
    psych::alpha(queer.construct[,44:51])$total$raw_alpha,
    psych::alpha(queer.construct[,31:51])$total$raw_alpha
  )
)

reliability_kable$Alpha <- round(reliability_kable$Alpha, 2)

reliability_kable <- reliability_kable %>%
  mutate(Interpretation = case_when(
    Alpha < 0.60 ~ "Poor",
    Alpha < 0.70 ~ "Questionable",
    Alpha < 0.80 ~ "Acceptable",
    Alpha < 0.90 ~ "Good",
    Alpha >= 0.90 ~ "Excellent"
  ))

kable(reliability_kable, 
      caption = "Reliability Tests Results",
      col.names = c("Scale", "Number of Items", "Cronbach's alpha", "Interpretation"),
      align = c("l", "c", "c", "c"),
      booktabs = TRUE,
      format = "latex") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = FALSE) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "2.5cm") %>%
  column_spec(3, width = "2.5cm") %>%
  column_spec(4, width = "2.5cm")

```

```{r}
#| label: Prepare for inline character display
#| include: false

sexual_orientation_inline<- c("asexual","bisexual","gay or lesbian","pansexual","and queer")

```

### Respondent Characteristics
167 respondents were eligible for the data analysis after data cleaning. Respondents ranged in age from `r min(queer.hypo$age)` to `r max(queer.hypo$age)` years old. The average age of the sample was `r round(mean(queer.hypo$age), 1)` years old (*SD* = `r round(sd_age, 1)`). This wide age range ensured the study captured perspectives from individuals across different generations. Respondents' residential countries included `r n_distinct (queer.hypo.country$country)` countries and four continents (Figure 1), with both developed and developing country status, providing a global perspective. Regarding sexuality, the sample featured a diverse range of identities, with `r round(100 * sum(queer.hypo$gender_identity == "Man") / nrow(queer.hypo), 1)`% identifying as men, `r round(100 * sum(queer.hypo$gender_identity == "Woman") / nrow(queer.hypo), 1)`% as women, and 10.2% as non-binary, genderfluid, gender non-conforming, or agender. Among the sample, `r round(100 * sum(queer.hypo$transgender == "Yes") / nrow(queer.hypo), 1)`% was identified as transgender. The sexual orientation distribution of participants was also diverse, featuring `r sexual_orientation_inline` (Figure 2). The racial distribution of the sample covered `r round(100 * sum(queer.hypo$race == "White") / nrow(queer.hypo), 1)`% white respondents, `r round(100 * sum(queer.hypo$race == "Black or African-American") / nrow(queer.hypo), 1)`% black or African-American respondents, and 15.5% Asian, Indian, or Alaska native. This rich diversity in age (generation), residential country, gender identity, sexual orientation, and racial identity enables meaningful comparisons and analyses across various LGBTQ+ subgroups and their intersecting identities.

::: {#fig-residential layout-ncol=2}

![Country Distribution](Image/Fig 01_Number of participants by country_bar.png){#fig-country}

![Continent Distribution](Image/Fig 02_Number of participants by country_map.png){#fig-continent}

Residential Distribution

:::

![Sexuality Distribution](Image/Fig 03_Gender identity_Transgender status_Sexual orientation_pie.png)

# Results

## Queer Media Characters’ Psychological Impacts on Different LGBTQ+ Subgroups

```{r}
#| label: ANOVA tests for psychological well-being scores by different LGBTQ+ subgroups
#| include: false

# 1. Age
anova_age_effect <- aov(psychological_wellbeing_score ~ generation, data = queer.hypo)
summary(anova_age_effect)

# p = .70, no significant difference has been found among different generations

# 2. Residential Country Development Status
anova_country_effect <- aov(psychological_wellbeing_score ~ development_status, data = queer.hypo)
summary(anova_country_effect)
summary_country_effect <- summary(anova_country_effect)[[1]]

# p = .014, F = 6.16, significant difference has been found between developed and developing countries

# 3. Gender identity
anova_gender_identity_effect <- aov(psychological_wellbeing_score ~ gender_identity_dual, data = queer.hypo)
summary (anova_gender_identity_effect)
summary_gender_effect <- summary(anova_gender_identity_effect)[[1]]

# p = .11, no significant difference has been found among different gender identity, but there was a directional effect

# 4. Transgender status
anova_transgender_effect <- aov(psychological_wellbeing_score ~ transgender, data = queer.hypo)
summary (anova_transgender_effect)

# p = .25, no significant difference has been found among different gender identity

# 5. Sexual orientation
anova_sexual_orientation_effect <- aov(psychological_wellbeing_score ~ sexual_orientation, data = queer.hypo)
summary (anova_sexual_orientation_effect)
summary_orientation_effect <- summary(anova_sexual_orientation_effect)[[1]]

# p = 0.1, no significant differences have been found among different sexual orientation, but there was directional effect

# 6. Race
anova_race_effect <- aov(psychological_wellbeing_score ~ race_category, data = queer.hypo)
summary(anova_race_effect)
summary_race_effect <- summary(anova_race_effect)[[1]]

# p = .09, marginally significant difference has been found between white and other

```

One-way ANOVA results showed that there were no significant differences or directional effects in queer media characters’ psychological impacts among different generations (*p* = .70) or between transgenders and non-transgenders (*p* = .25). However, significant differences were found between developed and developing countries' residents (Figure 3). Specifically, residents in developed countries showed lower psychological impacts scores by queer media characters than those in developing countries ($M_{\text{developed}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$development_status == "Developed"]), 2)`$ vs. $M_{\text{developing}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$development_status == "Developing"]), 2)`$; *F*(`r summary_country_effect$Df[1]`, `r summary_country_effect$Df[2]`) = `r round(summary_country_effect$"F value"[1], 2)`; *p* = `r sprintf("%.3f", summary_country_effect$"Pr(>F)"[1])`).

Though the ANOVA result did not show significant differences among different gender identities, there was a directional effect indicated by the violin plot (Figure 4). Specifically, men showed lower psychological impacts scores than women and other gender identities ($M_{\text{men}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$gender_identity_dual == "Man"]), 2)`$ vs. $M_{\text{womenother}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$gender_identity_dual == "Women, or nonbinary, or agender"]), 2)`$; *F*(`r summary_gender_effect$Df[1]`, `r summary_gender_effect$Df[2]`) = `r round(summary_gender_effect$"F value"[1], 2)`; *p* = `r sprintf("%.3f", summary_gender_effect$"Pr(>F)"[1])`). The distribution also indicated that men have more low scores than their women or nonbinary counterparts. Regarding sexual orientation, the ANOVA result did not show significant differences among different sexual orientations, but there was also a directional effect (Figure 5). Specifically, bisexual individuals showed the lowest psychological impacts scores among all sexual orientations, while gay or lesbian had the highest impacts scores (*F*(`r summary_orientation_effect$Df[1]`, `r summary_orientation_effect$Df[2]`) = `r round(summary_orientation_effect$"F value"[1], 2)`; *p* = `r sprintf("%.3f", summary_orientation_effect$"Pr(>F)"[1])`).

```{r}
#| label: fig-plot-country-psych-impact
#| fig-cap: Country Comparison of Queer Characters' Psychological Impacts
#| fig-width: 9
#| fig-height: 5

# Define the theme
theme.clean <- function(){
  theme_bw()+
    theme(axis.text.x = element_text(size = 12, vjust = 1, hjust = 1),
          axis.text.y = element_text(size = 12),
          axis.title.x = element_text(size = 14, face = "plain"),             
          axis.title.y = element_text(size = 14, face = "plain"),
          panel.grid.major.x = element_blank(),                                          
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),  
          plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
          plot.title = element_text(size = 16, vjust = 1, hjust = 0.5),
          plot.caption = element_text (size = 8,vjust = 1, hjust = 0.9),
          legend.text = element_text(size = 12, face = "italic"),          
          legend.position = "right")}

# Visualization: Boxplot

country_effect_boxplot <- ggplot(queer.hypo, aes(development_status, psychological_wellbeing_score)) + 
  geom_boxplot(aes(fill = development_status)) +
  theme_bw() +
  scale_fill_manual(values = c("#dbf1e5", "#4fba7f")) +               
  scale_colour_manual(values = c("#dbf1e5", "#4fba7f")) + 
  labs(title = "Country Comparison of Queer Characters' Psychological Impacts", 
       x = "Development Status", y = "Queer Characters' Psychological Impact Score") +
  theme.clean()+
  theme(axis.text.x = element_text(hjust = 0.5))+
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 3, shape = 18, 
               position = position_dodge(0.75), aes(shape = "Mean"), 
               show.legend = FALSE) +  
  stat_summary(fun = mean, geom = "text", aes(label = round(after_stat(y), 2)), 
               color = "black", vjust = -1, position = position_dodge(0.75), 
               size = 4, show.legend = FALSE) +
  labs(fill = "Development Status")

print (country_effect_boxplot)

```

```{r}
#| label: fig-plot-gender-psych-impact
#| fig-cap: Gender Identity Difference of Queer Characters' Psychological Impacts
#| fig-width: 10
#| fig-height: 5

# Visualization: Violinplot

gender_identity_effect_violinplot <- ggplot(queer.hypo, aes(gender_identity_dual, psychological_wellbeing_score)) + 
  geom_violin(aes(fill = gender_identity_dual)) +
  theme_bw() +
  scale_fill_manual(values = c("#6fa8dc", "#d5a6bd")) +               
  scale_colour_manual(values = c("#6fa8dc", "#d5a6bd")) + 
  labs(title = "Gender Identity Difference of Queer Characters' Psychological Impacts", 
       x = "Gender Identity", y = "Queer Characters' Psychological Impact Score") +
  theme.clean()+
  theme(axis.text.x = element_text(hjust = 0.5))+
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  stat_summary(fun = mean, geom = "point", color = "white", size = 3, shape = 18, 
               position = position_dodge(0.75), aes(shape = "Mean"), 
               show.legend = FALSE) +  
  stat_summary(fun = mean, geom = "text", aes(label = round(after_stat(y), 2)), 
               color = "white", vjust = -1, position = position_dodge(0.75), 
               size = 4, show.legend = FALSE) +
  labs(fill = "Gender Identity")

print (gender_identity_effect_violinplot)

```

![Sexual Orientation Comparison of Queer Characters' Psychological Impacts](Image/Fig 10_Orientation_Psych_box.png)

Racial differences in queer media characters’ psychological impacts were marginally significant (Figure 6). Specifically, white respondents showed lower psychological impacts scores than other racial groups ($M_{\text{white}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$race_category == "White"]), 2)`$ vs. $M_{\text{blackother}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$race_category == "Black, Asian, Indian, and Others"]), 2)`$; *F*(`r summary_race_effect$Df[1]`, `r summary_race_effect$Df[2]`) = `r round(summary_race_effect$"F value"[1], 2)`; *p* = `r sprintf("%.3f", summary_race_effect$"Pr(>F)"[1])`).

```{r}
#| label: fig-plot-race-psych-impact
#| fig-cap: Race Comparison of Queer Characters' Psychological Impacts
#| fig-width: 10
#| fig-height: 5
# Visualization: Violinplot
race_effect_boxplot <- ggplot(queer.hypo, aes(race_category, psychological_wellbeing_score)) + 
  geom_boxplot(aes(fill = race_category)) +
  theme_bw() +
  scale_fill_manual(values = c("#cfb5a0","#f1e2c8")) +               
  scale_colour_manual(values = c("#cfb5a0","#f1e2c8")) + 
  labs(title = "Race Comparison of Queer Characters' Psychological Impacts", 
       x = "Racial Identity", y = "Queer Characters' Psychological Impact Score") +
  theme.clean()+
  theme(axis.text.x = element_text(hjust = 0.5))+
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  stat_summary(fun = mean, geom = "point", color = "white", size = 3, shape = 18, 
               position = position_dodge(0.75), aes(shape = "Mean"), 
               show.legend = FALSE) +  
  stat_summary(fun = mean, geom = "text", aes(label = round(after_stat(y), 2)), 
               color = "white", vjust = -1, position = position_dodge(0.75), 
               size = 4, show.legend = FALSE) +
  labs(fill = "Racial Identity")

print (race_effect_boxplot)

```

## Predictors of Queer Media Characters’ Psychological Impacts

### Queer Media Consumption Habits

Data visualization of bubble plots revealed that queer media exposure frequency and engagement may predict queer media characters’ psychological impacts on LGBTQ+ individuals. The bubble plot for queer media exposure and discussion frequency indicated that developed countries have more participants who engaged less with queer media characters, which may contribute to lower impact scores (Figure 7). Though bubbles for developing countries were smaller, it was the result of a much smaller sample size (N = 22) for developing countries' residents than for developed countries' residents (N = 145). The bubble plot featuring different gender identities also showed that women, nonbinary, and agender respondents have higher discussion frequency of queer media representations than men (Figure 8), which may contribute to higher community belong scores and the overall psychological impact scores.

```{r}
#| label: fig-plot-country-freq
#| fig-cap: Country Differences in Queer Media Exposure Frequency and Engagement
#| fig-width: 15
#| fig-height: 5

create_country_bubble_plot <- function(queer.hypo, y_var, title) {
  ggplot(queer.hypo, aes(x = development_status, y = !!sym(y_var))) +
    geom_count(aes(size = ..n..), alpha = 0.6, color = "blue") +  
    labs(title = title,
         x = "Development Status",
         y = y_var) +
    scale_size_area(max_size = 15, name = "Count") +  
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5),
      axis.title.x = element_text(hjust = 0.5),  
      axis.text.x = element_text(hjust = 0.5),  
      legend.position = "bottom",  
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 7),  
      legend.key.size = unit(0.5, "cm"),
      plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")  # Reduce margin
    ) +
    guides(size = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5))}  # Adjust legend layout

country_freq_bubbleplot <- create_country_bubble_plot(queer.hypo, "frequency", "Frequency")
country_seek_bubbleplot <- create_country_bubble_plot(queer.hypo, "actively_seek_out", "Actively Seek Out Queer Representations")
country_discussion_bubbleplot <- create_country_bubble_plot(queer.hypo, "discussion", "Discussion Engagement of Queer Media Representations")

combined_country_bubble_plot <- grid.arrange(country_freq_bubbleplot, country_seek_bubbleplot, country_discussion_bubbleplot, ncol = 3)

```

```{r}
#| label: fig-plot-gender-freq
#| fig-cap: Gender Differences in Queer Media Exposure Frequency and Engagement
#| fig-width: 15
#| fig-height: 5

create_gender_identity_bubble_plot <- function(queer.hypo, y_var, title) {
  ggplot(queer.hypo, aes(x = gender_identity_dual, y = !!sym(y_var))) +
    geom_count(aes(size = ..n..), alpha = 0.6, color = "blue") +  
    labs(title = title,
         x = "Gender Identity",
         y = y_var) +
    scale_size_area(max_size = 15, name = "Count") +  
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5),
      axis.title.x = element_text(hjust = 0.5),  
      axis.text.x = element_text(hjust = 0.5),  
      legend.position = "bottom",  
      legend.title = element_text(size = 8),  
      legend.text = element_text(size = 7),  
      legend.key.size = unit(0.5, "cm"),
      plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")  
    ) +
    guides(size = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5))}

gender_identity_freq_bubbleplot <- create_gender_identity_bubble_plot(queer.hypo, "frequency", "Frequency")
gender_identity_seek_bubbleplot <- create_gender_identity_bubble_plot(queer.hypo, "actively_seek_out", "Actively Seek Out Queer Representations")
gender_identity_discussion_bubbleplot <- create_gender_identity_bubble_plot(queer.hypo, "discussion", "Discussion Engagement of Queer Media Representations")

combined_gender_identity_bubble_plot <- grid.arrange(gender_identity_freq_bubbleplot, gender_identity_seek_bubbleplot, gender_identity_discussion_bubbleplot, ncol = 3)

```

### Perceived Queer Character Authenticity

```{r}
#| label: Linear regression for authenticity rating predicting psychological well-being scores
#| include: false

authenticity.effect <- lm(psychological_wellbeing_score ~ authenticity_rating, data = queer.hypo)
summary(authenticity.effect)

summary_authenticity <- summary(authenticity.effect)

f_statistic_auth <- summary_authenticity$fstatistic[1]
df1_auth <- summary_authenticity$fstatistic[2]
df2_auth <- summary_authenticity$fstatistic[3]
r_squared_auth <- summary_authenticity$r.squared[1]
p_value_auth <- pf(f_statistic_auth, df1_auth, df2_auth, lower.tail = FALSE)

```

A linear regression was conducted to examine the relationship between queer character authenticity rating and psychological wellbeing impact scores. The results indicated a significant effect of authenticity rating on psychological wellbeing (*F*(`r df1_auth`, `r df2_auth`) = `r round(f_statistic_auth, 2)`, *p* `r if(p_value_auth < .001) "< .001" else paste("=", round(p_value_auth, 3))`). The $R^2$ was `r round(r_squared_auth,3)`, indicating that the perceived queer character authenticity explained `r round(r_squared_auth * 100, 1)`% of the variance in psychological wellbeing scores. This result suggests that the more authentic LGBTQ+ individuals perceive queer media characters, the more positive psychological impacts they may experience.

Figure 9 illustrates that the authencity rating positively predicts psychological well-being scores for both developed and developing countries' residents. The regression line for developing countries' residents is steeper than that for developed countries' residents, indicating that the authenticity rating has a stronger effect on psychological well-being scores for developing countries' residents.

Figure 10 indicates that the authenticity rating positively predicts psychological well-being scores for all gender identities. The regression line for men is steeper, indicating that the authenticity rating has a stronger effect on psychological wellbeing scores for men. Nevertheless, given the fact that men have lower psychological impacts scores, the authenticity rating may not be the only factor influencing psychological impact scores.

```{r}
#| label: fig-plot-auth-psych-impact-country
#| fig-cap: Authenticity Rating Predicting Psychological Well-being Scores by Country
#| message: false
#| fig-width: 10
#| fig-height: 5
auth.psych.dev <- ggplot(queer.hypo, aes(x = authenticity_rating, y = psychological_wellbeing_score)) +
  geom_point(aes(colour = development_status, shape = development_status)) +                                
  labs(x = "Authenticity Rating", y = "Queer Characters' Psychological Impact Score") +
  labs(title = "Authenticity Rating Predicting Psychological Effects by Development Status") + 
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  scale_x_continuous(
    breaks = seq(floor(min(queer.hypo$authenticity_rating)), 
                 ceiling(max(queer.hypo$authenticity_rating)), by = 1),
    limits = c(floor(min(queer.hypo$authenticity_rating)), ceiling(max(queer.hypo$authenticity_rating)))
  ) +
  stat_smooth(method = "lm", aes(fill = development_status, colour = development_status)) +    
  scale_colour_manual(name = "Development Status", values = c("#FFC125", "#36648B")) +
  scale_fill_manual(name = "Development Status",values = c("#FFC125", "#36648B")) +
  scale_shape_discrete(name = "Development Status") +
  theme.clean()

print(auth.psych.dev)

```

```{r}
#| label: fig-plot-auth-psych-impact-gender
#| fig-cap: Authenticity Rating Predicting Psychological Well-being Scores by Gender Identity
#| message: false
#| fig-width: 10
#| fig-height: 5.5

auth.psych.gender <- ggplot(queer.hypo, aes(x = authenticity_rating, y = psychological_wellbeing_score)) +
  geom_point(aes(colour = gender_identity_dual, shape = gender_identity_dual)) +                                
  labs(x = "Authenticity Rating", y = "Queer Characters' Psychological Impact Score") +
  labs(title = "Authenticity Rating Predicting Psychological Effects by Gender Identity") + 
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  scale_x_continuous(
    breaks = seq(floor(min(queer.hypo$authenticity_rating)), 
                 ceiling(max(queer.hypo$authenticity_rating)), by = 1),
    limits = c(floor(min(queer.hypo$authenticity_rating)), ceiling(max(queer.hypo$authenticity_rating)))
  ) +
  stat_smooth(method = "lm", aes(fill = gender_identity_dual, colour = gender_identity_dual)) +    
  scale_colour_manual(name = "Gender Identity", values = c("#6fa8dc", "#d5a6bd")) +
  scale_fill_manual(name = "Gender Identity", values = c("#6fa8dc", "#d5a6bd")) +
  scale_shape_discrete(name = "Gender Identity") +
  theme.clean()

print(auth.psych.gender)

```

Figure 11 illustrates that the authenticity rating positively predicts psychological well-being scores for all sexual orientations.It is worth noting that while data points for asexual, gay or lesbian, and pansexual concentrated on the middle to higher end of the authenticity rating and psychological well-being scores, data points for bisexual were more dispersed with much more low scores for both ratings. This result suggests that bisexual respondents' low psychological impacts scores may be related to the perceived authenticity of queer media characters.

Figure 12 shows that the authenticity rating positively predicts psychological well-being scores for all racial identities. The regression line for black, asian, and indian respondents is slightly steeper, indicating that racial minorities' perceived psychological impacts by queer characters may be more influenced by the perceived character authenticity than white respondents.

```{r}
#| label: fig-plot-auth-psych-impact-orientation
#| fig-cap: Authenticity Rating Predicting Psychological Well-being Scores by Sexual Orientation
#| message: false
#| fig-width: 11
#| fig-height: 5.5

auth.psych.orientation <- ggplot(queer.hypo_filtered, aes(x = authenticity_rating, y = psychological_wellbeing_score)) +
  geom_point(aes(colour = sexual_orientation, shape = sexual_orientation)) +                                
  labs(x = "Authenticity Rating", y = "Queer Characters' Psychological Impact Score") +
  labs(title = "Authenticity Rating Predicting Psychological Effects by Sexual Orientation") + 
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  scale_x_continuous(
    breaks = seq(floor(min(queer.hypo$authenticity_rating)), 
                 ceiling(max(queer.hypo$authenticity_rating)), by = 1),
    limits = c(floor(min(queer.hypo$authenticity_rating)), ceiling(max(queer.hypo$authenticity_rating)))
  ) +
  stat_smooth(method = "lm", aes(fill = sexual_orientation, colour = sexual_orientation)) +    
  scale_colour_manual(name = "Sexual Orientation", values = c("#A3A3A3","#9B4F96","#3D1A78","#FFD800")) +
  scale_fill_manual(name = "Sexual Orientation", values = c("#A3A3A3","#9B4F96","#3D1A78","#FFD800")) +
  scale_shape_discrete(name = "Sexual Orientation") +
  facet_wrap(~sexual_orientation) +
  theme.clean()

print(auth.psych.orientation)

```

```{r}
#| label: fig-plot-auth-psych-impact-race
#| fig-cap: Intersectionality Rating Predicting Psychological Well-being Scores by Racial Identity
#| message: false
#| fig-width: 11.5
#| fig-height: 5.5

auth.psych.race <- ggplot(queer.hypo, aes(x = authenticity_rating, y = psychological_wellbeing_score)) +
  geom_point(aes(colour = race_category, shape = race_category)) +                                
  labs(x = "Authenticity Rating", y = "Queer Characters' Psychological Impact Score") +
  labs(title = "Authenticity Rating Predicting Psychological Effects by Racial Identity") + 
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  scale_x_continuous(
    breaks = seq(floor(min(queer.hypo$authenticity_rating)), 
                 ceiling(max(queer.hypo$authenticity_rating)), by = 1),
    limits = c(floor(min(queer.hypo$authenticity_rating)), ceiling(max(queer.hypo$authenticity_rating)))
  ) +
  stat_smooth(method = "lm", aes(fill = race_category, colour = race_category)) +    
  scale_colour_manual(name = "Racial Identity", values = c("black", "#cfb5a0")) +
  scale_fill_manual(name = "Racial Identity", values = c("black","#cfb5a0")) +
  scale_shape_discrete(name = "Racial Identity") +
  theme.clean()

print(auth.psych.race)

```

### Perceived Queer Character Intersectionality

```{r}
#| label: Linear regression for intersectionality rating predicting psychological well-being scores
#| include: false

inter.effect <- lm(psychological_wellbeing_score ~ intersectionality_rating, data = queer.hypo)
summary(inter.effect)

summary_intersectionality <- summary(inter.effect)

f_statistic_inter <- summary_intersectionality$fstatistic[1]
df1_inter <- summary_intersectionality$fstatistic[2]
df2_inter <- summary_intersectionality$fstatistic[3]
r_squared_inter <- summary_intersectionality$r.squared[1]
p_value_inter <- pf(f_statistic_inter, df1_inter, df2_inter, lower.tail = FALSE)

```

Linear regression results also indicated a significant effect of queer character intersectionality rating on psychological wellbeing impact scores (*F*(`r df1_inter`, `r df2_inter`) = `r round(f_statistic_inter, 2)`, *p* `r if(p_value_inter < .001) "< .001" else paste("=", round(p_value_inter, 3))`). The $R^2$ was `r round(r_squared_inter,3)`, indicating that the perceived queer character authenticity explained `r round(r_squared_inter * 100, 1)`% of the variance in psychological wellbeing scores. This robust association suggests that LGBTQ+ individuals who perceive queer media characters as more intersectional tend to experience more positive psychological outcomes.

Figure 13 further indicated that queer characters' intersectionality may be a more significant contributor to psychological impact scores for developing countries' residents than for developed countries' residents. The regression line for developing countries' residents is steeper, with no intersectionality ratings lower than the middle point. Conversely, the regression line for developed countries' residents is flatter, with more low intersectionality ratings and psychological impact scores.

Though the regressioin line slopes for different gender identities were similar regarding intersectionality rating's effects on psychological impact scores, data points for women and other gender identities were more concentrated on the higher end, with lower intersectionality scores predicting relatively higher psychological impact scores than their men counterparts (Figure 14). Men respondents featured more low psychological impact data points even when the intersectionality rating was high. This indicates that the intersectionality rating may contribute to higher psychological impact scores for women and other gender identities, which could be a plausible explanation of men's lower psychological impact scores even if the authenticity ratings predict higher scores for men compared with women and other gender identities.

```{r}
#| label: fig-plot-inter-psych-impact-country
#| fig-cap: Authenticity Rating Predicting Psychological Well-being Scores by Country
#| message: false
#| fig-width: 10
#| fig-height: 5

inter.psych.dev <- ggplot(queer.hypo, aes(x = intersectionality_rating, y = psychological_wellbeing_score)) +
  geom_point(aes(colour = development_status, shape = development_status)) +                                
  labs(x = "Intersectionality Rating", y = "Queer Characters' Psychological Impact Score") +
  labs(title = "Intersectionality Rating Predicting Psychological Effects by Development Status") + 
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  scale_x_continuous(
    breaks = seq(floor(min(queer.hypo$intersectionality_rating)), 
                 ceiling(max(queer.hypo$intersectionality_rating)), by = 1),
    limits = c(floor(min(queer.hypo$intersectionality_rating)), ceiling(max(queer.hypo$intersectionality_rating)))
  ) +
  stat_smooth(method = "lm", aes(fill = development_status, colour = development_status)) +    
  scale_colour_manual(name = "Development Status", values = c("#FFC125", "#36648B")) +
  scale_fill_manual(name = "Development Status",values = c("#FFC125", "#36648B")) +
  scale_shape_discrete(name = "Development Status") +
  theme.clean()

print(inter.psych.dev)

```

```{r}
#| label: fig-plot-inter-psych-impact-gender
#| fig-cap: Intersectionality Rating Predicting Psychological Well-being Scores by Gender Identity
#| message: false
#| fig-width: 10
#| fig-height: 5.5

inter.psych.gender <- ggplot(queer.hypo, aes(x = intersectionality_rating, y = psychological_wellbeing_score)) +
  geom_point(aes(colour = gender_identity_dual, shape = gender_identity_dual)) +                                
  labs(x = "Intersectionality Rating", y = "Queer Characters' Psychological Impact Score") +
  labs(title = "Intersectionality Rating Predicting Psychological Effects by Gender Identity") + 
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  scale_x_continuous(
    breaks = seq(floor(min(queer.hypo$intersectionality_rating)), 
                 ceiling(max(queer.hypo$intersectionality_rating)), by = 1),
    limits = c(floor(min(queer.hypo$intersectionality_rating)), ceiling(max(queer.hypo$intersectionality_rating)))
  ) +
  stat_smooth(method = "lm", aes(fill = gender_identity_dual, colour = gender_identity_dual)) +    
  scale_colour_manual(name = "Gender Identity", values = c("#6fa8dc", "#d5a6bd")) +
  scale_fill_manual(name = "Gender Identity", values = c("#6fa8dc", "#d5a6bd")) +
  scale_shape_discrete(name = "Gender Identity") +
  theme.clean()

print(inter.psych.gender)

```

Based on the finding that intersectionality rating contributed to higher psychological impact scores for women and other gender identities, Figure 15 compared both gender identity and sexual orientation on differences in intersectionality rating's prediction of psychological impact scores. The results indicated that intersectionality rating was a stronger predictor of psychological impact scores for bisexual men compared with bisexual women and other gender identities, with a steeper regression line. Nevertheless, the regression line for lesbian women is slightly steeper than gay men, indicating that intersectionality rating may have a stronger effect on psychological impact scores for lesbian than gay. It is worth noting that while the associations between intersectionality rating and psychological impact scores were positive among asexual and pansexual women with relatively small variations, the associations were negative among asexual and pansexual men with more variations and data point proportion below the regression line. Thus, it is plausiable that the negative associations contributed significantly to the lower psychological impact scores predicted by the intersectionality rating for men respondents.

```{r}
#| label: fig-plot-inter-psych-impact-orientation-gender
#| fig-cap: Intersectionality Rating Predicting Psychological Well-being Scores by Sexual Orientation and Gender Identity
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 5.5

inter.psych.orientation.gender <- ggplot(queer.hypo_filtered, aes(x = intersectionality_rating, y = psychological_wellbeing_score)) +
  geom_point(aes(colour = sexual_orientation, shape = sexual_orientation)) +                                
  labs(x = "Intersectionality Rating", y = "Queer Characters' Psychological Impact Score") +
  labs(title = "Intersectionality Rating Predicting Psychological Effects by Sexual Orientation and Gender Identity") + 
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  scale_x_continuous(
    breaks = seq(floor(min(queer.hypo$intersectionality_rating)), 
                 ceiling(max(queer.hypo$intersectionality_rating)), by = 1),
    limits = c(floor(min(queer.hypo$intersectionality_rating)), ceiling(max(queer.hypo$intersectionality_rating)))
  ) +
  stat_smooth(method = "lm", aes(fill = sexual_orientation, colour = sexual_orientation)) +    
  scale_colour_manual(name = "Sexual Orientation", values = c("#A3A3A3","#9B4F96","#3D1A78","#FFD800")) +
  scale_fill_manual(name = "Sexual Orientation", values = c("#A3A3A3","#9B4F96","#3D1A78","#FFD800")) +
  scale_shape_discrete(name = "Sexual Orientation") +
  facet_grid(gender_identity_dual ~ sexual_orientation) +
  theme.clean() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 8),
        legend.position = "bottom")

print(inter.psych.orientation.gender)

```

Figure 16 illustrates that the intersectionality rating positively predicts psychological well-being scores for all racial identities. The regression line for black, asian, and indian respondents is steeper with more data points concentrated on the higher end of the intersectionality rating and psychological impact scores. This indicates that racial minorities' perceived psychological impacts by queer characters may be more influenced by the perceived character intersectionality than white respondents.

```{r}
#| label: fig-plot-inter-psych-impact-race
#| fig-cap: Intersectionality Rating Predicting Psychological Well-being Scores by Racial Identity
#| message: false
#| fig-width: 11.5
#| fig-height: 5.5

inter.psych.race <- ggplot(queer.hypo, aes(x = intersectionality_rating, y = psychological_wellbeing_score)) +
  geom_point(aes(colour = race_category, shape = race_category)) +                                
  labs(x = "Intersectionality Rating", y = "Queer Characters' Psychological Impact Score") +
  labs(title = "Intersectionality Rating Predicting Psychological Effects by Racial Identity") + 
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  scale_x_continuous(
    breaks = seq(floor(min(queer.hypo$intersectionality_rating)), 
                 ceiling(max(queer.hypo$intersectionality_rating)), by = 1),
    limits = c(floor(min(queer.hypo$intersectionality_rating)), ceiling(max(queer.hypo$intersectionality_rating)))
  ) +
  stat_smooth(method = "lm", aes(fill = race_category, colour = race_category)) +    
  scale_colour_manual(name = "Racial Identity", values = c("black", "#cfb5a0")) +
  scale_fill_manual(name = "Racial Identity", values = c("black", "#cfb5a0")) +
  scale_shape_discrete(name = "Racial Identity") +
  theme.clean()

print(inter.psych.race)

```

# Discussion

## Limitations and Future Directions

## Conclusion

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

# Appendix

# Title for Appendix